A Survey on Ordered Binary Decision Diagrams

### 2015-09-15 paper.tex
Revise BEDs

In 1997, Andersen and Hulgaard~\cite{ANDERSEN97} proposeded extending OBDDs with
a third type of nodes representing \textit{operators}. The resulting
representation is referred to as \textit{Binary Expression Diagram}
(\textit{BED}). Such an operator node $n$ has one attribute $op(n)$, the binary
Boolean operator, and, like nonterminal nodes, two children $lo(n)$ and $hi(n)$.
Operator nodes represent their corresponding Boolean connectives in the BED
graph~\cite{ANDERSEN97}.

Just like OBDDs, BEDs share equivalent Boolean subexpressions. Furthermore, BEDs
have their own set of manipulating operations, especially because of the operator
nodes. For example, the \textit{up-step} operation moves variables up in the BED
hierarchy, if possible. As an effect, operator nodes are pushed towards the
terminal nodes and some operator nodes may be evaluated if some of the associated
expressions are terminal nodes. The BED may be simplified afterwards. These steps
can be repeated to reduce the BED over and over again, and simplifying some tests,
e.g. tautology testing. It is also possible to move all variable up in a single
step. Nevertheless, reducing the BED step-by-step can exploit structural
information~\cite{ANDERSEN97}.

In contrast to OBDDs, BEDs are \textit{exponentially more succinct} because
arbitrary circuits can be translated into BEDs. As a consequence, the BED
representation suffices to solve integer multiplication using a quadratic number
of nodes~\cite{ANDERSEN97}.

Analogous to OBDDs and FBDDs, OBEDs and FBEDs can be defined, whereas OBEDs are
OBDDs without operator nodes. The reduction rule of BEDs is similar to the one
of OBDDs but extended with an additional rule for operator nodes. For any operator
nodes $n$ it has to hold that $lo(n)$ and $hi(n)$ are nonterminals in order to
guarantee reducedness. Apart from that, before generating a new node, an existence
check is performed. If there exists an equivalent node, this node is reused.
Otherwise the new node is created and considered in the next check. However, even
if a BED is reduced, canonicity is not guaranteed (in contrast to
OBDDs)~\cite{ANDERSEN97}.

Andersen and Hulgaard~\cite{ANDERSEN97} also introduced some checks especially
for operator nodes in order to reduce the BED-size even further. Moreover a
detailed description of how to convert a BED into an equivalent OBDD is provided,
distinguishing between step-by-step variable moving (denoted as \textit{up\_one})
and block-wise variable moving (denoted as \textit{up\_all}).

The \textit{up\_one} operation can be implemented with expected runtime of $O(n)$,
where $n$ denotes the number reachable of nodes from a given node. Nonetheless,
because BEDs are exponentially more succinct than OBDDs, the complete
transformation has an exponential worst-case complexity. Also \textit{up\_all} has
an exponential worst-case complexity (for the same reason)~\cite{ANDERSEN97}.

Andersen and Hulgaard~\cite{ANDERSEN97} concluded that \textit{up\_all} is a
memory efficient approach to construct OBDDs in contrast to the OBDD construction
from scratch, without actually increasing the runtime. \textit{up\_one} is used
to implement existential quantification and substitution efficiently, having
linear runtime complexity in the size of the BED. Reconsidering symbolic model
checking, BEDs can be used to evaluate fixed point computations directly without
converting it to an OBDD. Moreover, BEDs can perform fixed point computations more
efficiently because of their succinct representation~\cite{ANDERSEN97}. \tbr

>>>>

In 1997, Andersen and Hulgaard~\cite{ANDERSEN97} proposed to extend OBDDs with
a third type of nodes representing \textit{operators}. The resulting
representation is referred to as \textit{Binary Expression Diagram}
(\textit{BED}). Such an operator node $n$ has one attribute $op(n)$, the binary
Boolean operator, and, like nonterminal nodes, two children $lo(n)$ and $hi(n)$.
Operator nodes represent their corresponding Boolean connectives in the BED
graph~\cite{ANDERSEN97}.

Just like OBDDs, BEDs share equivalent Boolean subexpressions. Furthermore, BEDs
have their own set of manipulating operations, especially because of the operator
nodes. For example, the \textit{up-step} operation moves variables up in the BED
hierarchy, if possible. As an effect, operator nodes are pushed towards the
terminal nodes and some operator nodes may be evaluated if some of the associated
expressions are terminal nodes. The BED may be simplified afterwards. These steps
can be repeated to reduce the BED over and over again, and thus to simplify some
tests, e.g. tautology testing. It is also possible to move all variables up in a
single step. Nevertheless, reducing the BED step-by-step can exploit structural
information~\cite{ANDERSEN97}.

In contrast to OBDDs, BEDs are \textit{exponentially more succinct}, and arbitrary
circuits can be translated into BEDs. As a consequence, the BED representation
suffices to solve integer multiplication using a quadratic number of
nodes~\cite{ANDERSEN97}.

Analogous to OBDDs and FBDDs, OBEDs and FBEDs can be defined, whereas OBEDs are
OBDDs without operator nodes. The reduction rules of BEDs are similar to the rules
of OBDDs but extended with an additional rule for operator nodes. For any operator
node $n$ it has to hold that $lo(n)$ and $hi(n)$ are nonterminals in order to
guarantee reducedness. Apart from that, before generating a new node, an existence
check is performed. If there exists an equivalent node, this node is recycled.
Otherwise, the new node is created and considered in the next check. However, in
contrast to OBDDs, reducedness of BEDs does not imply
canonicity~\cite{ANDERSEN97}.

Andersen and Hulgaard~\cite{ANDERSEN97} also introduced some checks especially
for operator nodes in order to reduce the BED-size even further. Moreover, a
detailed description of how to convert a BED into an equivalent OBDD is provided,
distinguishing between step-by-step variable moving (denoted as \textit{up\_one})
and block-wise variable moving (denoted as \textit{up\_all}).

The \textit{up\_one} operation can be implemented with expected runtime of $O(n)$,
where $n$ denotes the number of reachable nodes from a given node. Nonetheless,
because BEDs are exponentially more succinct than OBDDs, the complete
transformation has an exponential worst-case complexity. Also \textit{up\_all} has
an exponential worst-case complexity (for the same reason)~\cite{ANDERSEN97}.

Andersen and Hulgaard~\cite{ANDERSEN97} concluded that \textit{up\_all} is a
memory efficient alternative to construct OBDDs compared to the OBDD construction
from scratch using \textit{Apply}, without actually increasing the runtime.
\textit{up\_one} is used to implement existential quantification and substitution
efficiently, having linear runtime complexity in the size of the BED.
Reconsidering symbolic model checking, BEDs can be used to evaluate fixed point
computations directly without converting to OBDDs. Moreover, BEDs can perform
fixed point computations more efficiently because of their succinct
representation~\cite{ANDERSEN97}.
